Unifying quasars with clumpy wind models

### 2015-07-31 paper.tex
few more mods

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:uvspec}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.png}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:lobal}
\end{figure}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in wuasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. Our \civ\ line emission region 
is located at about $1000~R_G$ ($\sim10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

Figure 4 shows the synthetic spectrum in the UV from the two models, with the
benchmark model from H13 also shown for comparison. We also show a comparison
to composite quasar and BALQSO spectra in figure 5, with the optical waveband 
included. In the UV, strong BAL profiles are formed in both models. This
highlights the first success of our model: clumping allows the correct ionization state 
to be maintained in the presence of strong X-rays. 
The BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At lower inclinations, the strongest absorption occurs at the red edge,
whereas at high inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.
% the line profile shape. Figure ?? shows selection of line profiles 
% for different kinematic parameters to demonstrate this. 
% The degeneracy between parameters, and lack of available
% physical and observational constraints goes someway to showing why individual 
% fitting of line profile shapes with kinematic models is such a difficult
% exercise.

@@

To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
The model matches the \civ\ and \mg\ line profiles fairly well, 
but the \la\ profile is a factor of $\sim3$ weaker than in the composite spectrum.
This is implies that a larger recombination line formation region is required in order 
to successfully reproduce the strong \la\ emission seen in 
all quasars. At the highest inclinations, the 
cooler, lower ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines lower ionization species such as \mg,
\al and Fe~\textsc{ii}, as well as strong Lyman continuum absorption.
The potential links to LoBALQSOs and FeLoBALQSOs are discussed in section 2.4.
% Figure 5 also shows the spectrum into the optical,
% showing that the outflow produces signiciant optical emission lines
% as well as having a noticeable effect on the continuum shape. This
% is particularly true around the Balmer jump at $3646\AA$ and is due to
% a prominent wind-formed recombination continuum.

@@

If the majority of continuum radiation was reprocessed by nearby photoionized plasma (such 
as the outflow itself), then this would result in a roughly
isotropic radiation field. There are a number of appealing aspects to this scenario.
First, it may go some way to explaining recent microlensing (REFs) and timing (REFs)
observations which suggest that accretion disc sizes are a factor of $\sim3$ 
larger than one expects from a Shakura \& Sunyaev (1973) thin disc. 
Second, it implies a wide angle, large covering factor wind in order to 
intercept most of the disc continuum. This would make the outflow 
an effective feedback mechanism \citep[e.g.][]{borguet2012}, 
and would imply that BALQSO outflows
have a similar geometry to the ultra-fast outflows seen in nearby AGN 
\citep[e.g.][]{reeves2003,poundsreeves2009,tombesi2010a}.
We note that spectropolarimetry of BALQSOs shows that the light from BAL troughs is $\sim10\%$ 
linearly polarised, but the continuum generally has negligible polarisation \citep{lamy2000}.
This suggests that the continuum of BALQSOs cannot be significantly enhanced by
{\em single} scattering into the line of sight, but does not rule out 
{\em multiple} scatterings leading to a more isotropic continuum.

>>>>

\begin{figure*} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/wind.png}
\caption
{
Physical properties of the outflow, shown by the coloured contours.
The solid black line marks a sphere at $1000~R_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
}
\label{fig:wind}
\end{figure*} %fullpage

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/sightlines.png}
\caption
{
The classes of sightlines and the corresponding possible 
system types in the framework of our model.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in wuasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

Clumping also causes the total line luminosity to increase dramatically,
as recombination and collisional excitation are both proportional to
$n_e^2$. This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the X-rays heating the plasma.
The radii of these emitting regions is important,
and can be compared to observations. As shown in ~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~R_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

Figure 4 shows the synthetic spectrum in the UV from the two models, with the
benchmark model from H13 also shown for comparison. We also show a comparison
to composite quasar and BALQSO spectra in figure 5, with the optical waveband 
included. In the UV, strong BAL profiles are formed in both models. This
highlights the first success of our model: clumping means the correct ionization state 
is maintained in the presence of strong X-rays, allowing large resonance line opacities. 
At the highest inclinations, the 
cooler, lower ionization material at the base of the wind
starts to intersect the line of sight. This produces 
multiple absorption lines lower ionization species such as \mg,
\al\ and Fe~\textsc{ii}, as well as strong Lyman continuum absorption.
The potential links to LoBALQSOs and FeLoBALQSOs are discussed in section 2.4.

The BAL profiles are often saturated, and the location in velocity space
of the strongest absorption in the profile varies with inclination.
At lower inclinations, the strongest absorption occurs at the red edge,
whereas at high inclinations (and for the strongest BALs)
the trough has a sharp edge at the terminal velocity.
This offers one potential explanation for the wide range of BALQSO absorption
line shapes (see e.g. Trump et al. 2006; Knigge et al 2008).
In addition, the line profile shape is strongly dependent 
on the density, ionization and velocity 
profiles intersected by the line of sight. Thus, small tweaks of the velocity
law and angular distributions of streamlines can dramatically alter
the shape of the line.
% the line profile shape. Figure ?? shows selection of line profiles 
% for different kinematic parameters to demonstrate this. 
% The degeneracy between parameters, and lack of available
% physical and observational constraints goes someway to showing why individual 
% fitting of line profile shapes with kinematic models is such a difficult
% exercise.

@@

To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar composites from
\cite{reichard2003}, normalised to the flux at 2000\AA\ in each panel.
The model matches the \civ\ and \mg\ line profiles fairly well, 
but the \la\ profile is a factor of $\sim3$ weaker than in the composite spectrum.
This is implies that a larger recombination line formation region is required in order 
to successfully reproduce the strong \la\ emission seen in 
all quasars. We do not produce the semi-forbidden 
intercombination lines seen in quasar spectra because we
currently do not have a treatment for semi-forbidden lines.
This is especially noteiceable with the strong [C~\textsc{iii}~1909\AA\, 
line in the quasar composite spectra.
The critical density of the [C~\textsc{iii}~1909\AA\, line 
is $n_e\sim10^9.5~\rm{cm^{-3}}$, which is higher than much of the 
outer portion of our wind. We therefore expect intercombination lines
to become important coolants in the outer portion of the wind.

% Figure 5 also shows the spectrum into the optical,
% showing that the outflow produces signiciant optical emission lines
% as well as having a noticeable effect on the continuum shape. This
% is particularly true around the Balmer jump at $3646\AA$ and is due to
% a prominent wind-formed recombination continuum.

@@

If the majority of continuum radiation was reprocessed by nearby photoionized plasma (such 
as the outflow itself), then this would result in a roughly
isotropic radiation field. There are a number of appealing aspects to this scenario.
First, it may go some way to explaining recent microlensing (REFs) and timing (REFs)
observations which suggest that accretion disc sizes are a factor of $\sim3$ 
larger than one expects from a Shakura \& Sunyaev (1973) thin disc. 
Second, it implies a wide angle, large covering factor wind in order to 
intercept most of the disc continuum. This would make the outflow 
an effective feedback mechanism \citep[e.g.][]{borguet2012}, 
and would imply that BALQSO outflows
have a similar geometry to the ultra-fast outflows seen in nearby AGN 
\citep{reeves2003,poundsreeves2009,tombesi2010a}.
We note that spectropolarimetry of BALQSOs shows that the light from BAL troughs is $\sim10\%$ 
linearly polarised, but the continuum generally has negligible polarisation \citep{lamy2000}.
This suggests that the continuum of BALQSOs cannot be significantly enhanced by
{\em single} scattering into the line of sight, but does not rule out 
{\em multiple} scatterings leading to a more isotropic continuum.
